---
title: "BIOL432 Assignment 1 Report"
author: "Sam Boots"
date: "January 13th 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[GitHub repository for BIOL432 Assignment 1](https://github.com/SBootss/BIOL432_Assignment1)

# Setup

## Loading libraries
Loading libraries for use in assignment:
```{r}
library(dplyr)
library(ggplot2)
```
## Loading default plotting theme
```{r}
source("http://bit.ly/theme_pub")
theme_set(theme_pub())
```

# Generating data and estimating limb volume
## R Script #1
The folowing script creates a data frame of 100 randomly sampled values for each of the variables organism, limb-width, limb-length and observer. The data frame is then exported to a csv called 'measurements.csv'. 
```{r}
source('dataGenerato.R')
```

## R Script #2
The following script adds a "Volume" column, estimating limb-volume from limb-width and limb-length with the assumption that the organism's limbs are cylindrical. The script overrides the original measurements.csv file with the new one including volume estimates.
```{r}
source('volumeEstimato.R')
```

# Loading data
Importing limb volume data from measurements.csv:

```{r}
LimbData<-read.csv("measurements.csv")
```

# Sorting data using dplyr

By species:

```{r}
LimbData %>%
  arrange(Organism)
```
By observer:

```{r}
LimbData %>%
  arrange(Observer)
```
By limb volume:

```{r}
LimbData %>%
  arrange(Volume)
```
# Creating summary tables
A table showing average limb-volume by species:
```{r}
LimbVolSpecAvg<-LimbData %>%
  group_by(Organism) %>%
  summarise("Average Limb Volume (cm^3)"=mean(Volume))

LimbVolSpecAvg
```
A table showing number of observations for each combination of species and observer:
```{r}
LimbVolSpObCount<-LimbData %>%
  count(Organism, Observer)

LimbVolSpObCount
```
# Figures summarizing limb volume data

## Box plot comparing limb volumes between species
```{r}
VolBoxPlot<-ggplot(LimbData, aes(Organism, Volume, fill=Organism)) +
  geom_boxplot()+
  labs(x="Species", y=expression(Limb~Volume~(cm^3)))+ scale_y_continuous(limits=c(0,5000),breaks= seq(0,5000, by=500))+
   theme(axis.ticks=element_line(), axis.text.x=element_text(face="italic"))
  

VolBoxPlot
```

## Frequency histograms of limb volume for each species

```{r}
FreqHist<-ggplot(LimbData, aes(x=Volume, fill=Organism))+
  geom_histogram(binwidth=500)+
  labs(x=expression(Limb~Volume~(cm^3)),y="Frequency")+
  scale_y_continuous(limits=c(0,10),
                     breaks=seq(0,10, by=2))+
  theme(axis.ticks=element_line(), strip.text=element_text(face="italic"))+
  facet_grid(.~Organism)

FreqHist
```

